### 2018-01-18 manuscript.rmd
modified conclusions, legends

![(A) Maximum likelihood (ML) phylogram with bipartition frequencies from the ML bootstraps and the Bayesian posterior distribution of trees. Unlabeled nodes have support >0.99 for both bootstraps and posteriors. (B) The topologies found in the posterior distribution of trees that conflict with the ML tree. (C) The topologies evaluated by the SOWH tests. For more details on the SOWH topologies refer to Fig. S2.](figures/Figure3.pdf)

@@

![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)

@@

How we define absence of tentilla, especially for *Physalia physalis*, is also important. The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@BARDI2007]. Siphonophore tentilla are complete diverticular branchings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).  *Physalia*'s buttons enclose individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm [@Hessinger1988]. This suggests they are not just ectodermal swellings, but probably reduced tentilla. When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig. S4), and in *Bathyphysa conifera*.

@@

Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. Using 33 siphonophore species, sampled across the phylogeny, we were able to gain more resolution of the major clades within Siphonophora. We suggest that monoecy likely evolved twice and siphonophores evolved very different ways to be monoecious, however we are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora. This phylogeny suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae and Calycophorae to further resolve the internal relationships within these clades.  The vast majority of valid described siphonophore species likely fall within these two groups, with 109 species in Calycophorae and 62 species in Euphysonectae.

>>>>

![(A) Maximum likelihood (ML) phylogram with bipartition frequencies from the ML bootstraps and the Bayesian posterior distribution of trees. Unlabeled nodes have support >0.99 for both bootstraps and posteriors. The numbers of valid described species estimated to be based in each clade based on taxonomy are shown below beach clade name on the right. (B) The topologies found in the posterior distribution of trees that conflict with the ML tree. (C) The topologies evaluated by the SOWH tests. For more details on the SOWH topologies refer to Fig. S2.](figures/Figure3.pdf)

@@

![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species. Bottom: siphonophore phylogeny, colored by clade. Middle panel: diagram showing the presence/absence of traits across Siphonophora, with the physical location of the trait shown on a schematic of *Nanomia bijuga* (schematic by Freya Goetz). Top: Bathymetric distribution of siphonophore species. Physalia illustration by Noah Schlottman, taken from http://phylopic.org/](figures/Figure4.pdf)

@@

How we define absence of tentilla, especially for *Physalia physalis*, is also important. The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@BARDI2007]. Siphonophore tentilla are complete diverticular branchings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).  *Physalia*'s buttons enclose individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm [@Hessinger1988]. This suggests they are not just ectodermal swellings, but probably reduced tentilla. When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig. S4), and in *Bathyphysa conifera*. The application of phylogenetic methods to the evolution of tentillum morphology would be a crucial step towards understanding the evolution of these structures, and their relationship with the feeding ecology of siphonophores.

@@

Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. Among other relationships, we identify the clade Euphysonectae as the sister group to Calycophorae. Our results suggest that monoecy arose twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.  We are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora, which will require greater taxon sampling and improved morphological understanding of many poorly known species. The improved resolution presented in this study suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae (where we sampled 13 of 62 valid described species that likely belong to the group) and Calycophorae (where we sampled 9 species in a clade of 109 valid described species) to further resolve the internal relationships within these clades.
